---
title: "Data Cleaning: Interfamilial Violence and Corruption in Guatemala"
author: "Rachel Rosenberg"
date: "7/10/2020"
output: 
  pdf_document: 
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setting up the Data

The [2012](https://docs.google.com/spreadsheets/d/1SfONash2AcxWDhOzdOZrOGsJhGOZq4XWeXJUzzOGvlQ/edit#gid=0) and [2013](https://docs.google.com/spreadsheets/d/1VDXSHcyIVZMQam1wqsf3_6ecSY8RI3PmudGR9AnNPlk/edit#gid=0) data sets include municipal-level information on birth rates, crime rates, homicide rates, and interfamiliar violence (i.e. domestic violence). 

**To begin:**

*Set the working directory and clear the environment;*
```{r}
#Set the working directory
setwd("/Users/racheleryn/Documents/Git/GuatemalaProject")

#Clear the environment
rm(list=ls(all=TRUE))
```

*Upload the 2012 and 2013 data;*
```{r}
#Call the rio package
library(rio) 

#Import the Excel documents
g_2012 = import("2012.xlsx") #the data from 2012 will be called g_2012
g_2013 = import("2013.xlsx") #the data from 2013 will be called g_2013 
```

*Preview the g_2012 data;*
```{r}
str(g_2012)
```

*Preview the g_2013 data;*
```{r}
str(g_2013)
```


## Section 1: Clean Data 
1. Put the municipality and department variables in lowercase with the (tolower) command.

  **For 2012:**
```{r}
#Lower case for the municipality and department variables for g_2012 data
g_2012$municipio = tolower(g_2012$municipio)
g_2012$departamento = tolower(g_2012$departamento)

#Check that the changes went through
head(g_2012, n = 1)
```

  **For 2013:**
```{r}
#Lowering case for the municipality and department variables for g_2013 data
g_2013$municipio = tolower(g_2013$municipio)
g_2013$departamento = tolower(g_2013$departamento)

#Check that the changes went through
head(g_2013, n = 1)
```

2. Remove the accents from those same lowercase department and municipality variables.
```{r}
#Create a function to remove accents 
remove.accents = function(s){
  old1 = "áéóíúñ"
  new1 = "aeoiun"
  s1 = chartr(old1, new1, s) 
}
```

  **For 2012:**
```{r}
#Remove accents for the g_2012 data
g_2012$departamento = remove.accents(g_2012$departamento)
g_2012$municipio = remove.accents(g_2012$municipio)

#Check that the changes went through
head(g_2012, n = 1)
```

  **For 2013:**
```{r}
#Remove accents for the g_2013 data
g_2013$departamento = remove.accents(g_2013$departamento)
g_2013$municipio = remove.accents(g_2013$municipio)

#Check that the changes went through
head(g_2013, n = 1)
```


3. Rename the variables so that there are no spaces and replace those spaces with underscores.

  **For 2012:**
```{r}
#Fix the column names for g_2012
colnames(g_2012) = gsub(" ", "_", colnames(g_2012))

#Check that the changes went through
head(g_2012, n = 1)
```

  **For 2013**    
```{r}
#Fix the column names for g_2013
colnames(g_2013) = gsub(" ", "_", colnames(g_2013))

#Check that the changes went through
head(g_2013, n = 1)
```


4. Add year variables to each of the files.
```{r, include = FALSE}
#Call the dplyr package
library(dplyr)
```
```{r}
#Create variables with the respective years
year12 = 2012
year13 = 2013

#Add the variables to the data frames using the dplyr package
g_2012 = mutate(g_2012,year12) 
g_2013 = mutate(g_2013,year13)

#Rename the variables using the dplyr package 
g_2012 = g_2012 %>% rename(year = year12)
g_2013 = g_2013 %>% rename(year = year13)

#Check that the changes went through
colnames(g_2012) #for 2012
colnames(g_2013) #for 2013
```

5. Ensure that all variables are the correct class.
```{r}
#Check the structure of the g_2012 data set
str(g_2012) 

#Check the structure of the g_2013 data set
str(g_2013) 
```

6. Label all of the variables using the labelled package.
```{r, include = FALSE}
library(labelled)
```
```{r}
#Change Labels for the g_2012 data set with the labelled package
var_label(g_2012) <- list(`departamento` = "Department",
                          `municipio` = "Municipality",
                          `tasa_bruta_de_natalidad` = "Birth Rate",
                          `tasa_de_criminalidad` = "Crime Rate",
                          `tasa_de_homicidios` = "Homicide Rate",
                          `violencia_intrafamiliar` = "Domestic Violence",
                          `year` = "Year")

#Change labels for the g_2013 data set
var_label(g_2013) <- list(`departamento` = "Department",
                          `municipio` = "Municipality",
                          `tasa_bruta_de_natalidad` = "Birth Rate",
                          `tasa_de_criminalidad` = "Crime Rate",
                          `tasa_de_homicidios` = "Homicide Rate",
                          `violencia_intrafamiliar` = "Domestic Violence",
                          `year` = "Year")
```


7. Save each of the cleaned cross-sectional data sets as Stata data sets.
```{r, eval = FALSE}
export(g_2012, file = "clean_g_2012.dta") #the clean data set for the g_2012 data
export(g_2013, file = "clean_g_2013.dta") #the clean data set for the g_2013 data
```


## Section 2: Append Panel Data
```{r, include = FALSE}
#Append g_2012 and g_2013
library(tidyverse)
```
```{r}
#Append g_2012 and g_2013
append_data = bind_rows(g_2012,g_2013)

#View the dimensions to check the append
dim(append_data)
```
```{r}
#Save the panel data in a Stata file
export(append_data, file = "panel_data.dta") #the appended 2012 and 2013 data frame
```

## Section 3: Merge Data Sets
```{r}
#Import the v5_guatemala_clean.dta file
v5_guatemala_cleaned = import("v5_guatemala_cleaned.dta") 

#Look at the structure of the data
dim(v5_guatemala_cleaned)
str(v5_guatemala_cleaned)
```

**Rename `municipio` and `departamento` columns in `append_data` to ensure they merge as the same columns in the `v5_guatemala_clean` data:**

```{r}
#Rename the variables using the dplyr package
append_data = append_data %>%
  rename(department = departamento,
         municipality = municipio)

#Check to make sure the changes went through
colnames(append_data)

#Adjust the labels with the labelled package
var_label(append_data) <- list(`department` = "Department",
                          `municipality` = "Municipality",
                          `tasa_bruta_de_natalidad` = "Birth Rate",
                          `tasa_de_criminalidad` = "Crime Rate",
                          `tasa_de_homicidios` = "Homicide Rate",
                          `violencia_intrafamiliar` = "Domestic Violence",
                          `year` = "Year")
```

**Merge the data:**

```{r}
#merge append_data and v5_guatemala_clean data 
v6_Guatemala = left_join(x = v5_guatemala_cleaned, y = append_data)

#check the merge
dim(v6_Guatemala)
colnames(v6_Guatemala)

#Save the new data set as a Stata file
export(v6_Guatemala, file = "v6_Guatemala.dta")
```
## Section 4: Prepare Specifications

**Specification context**

Domestic violence is the outcome that I'm interested in analyzing. I consider domestic violence the "x" variable in my future analyses. I consider corruption, mayor gender, crime rate, and homicide rate as potential "y" variables that may explain variation in "x" values (i.e. domestic violence). Note: This data set only includes domestic violence, homicide rates, and crime rates for years 2012 and 2013. I hope to add more years of these municipal-level data in the future.

For quick reference: `violencia_intrafamiliar` is the variable to account for domestic violence. `infrac_amount_` is the variable to account for corruption. `tasa_de_homicidios` is the variable to account for homicide rates. `tasa_de_criminalidad` is the variable to account for crime rates. `female` is a dummy variable to account for mayor gender.

```{r}
#Specification 1: corruption and domestic violence
specification1 = v6_Guatemala %>% drop_na("violencia_intrafamiliar", 
                                          "infrac_amount_")
dim(specification1)

#Specification 2: mayor gender and domestic violence
specification2 = v6_Guatemala %>% drop_na("violencia_intrafamiliar",
                                          "female")
dim(specification2)

#Specification 3: corruption, mayor gender, and domestic violence
specification3 = v6_Guatemala %>% drop_na("infrac_amount_", 
                                          "female", "violencia_intrafamiliar")
dim(specification3)

#Specification 4: crime rate, homicide rate, and domestic violence
specification4 = v6_Guatemala %>% drop_na("tasa_de_criminalidad", 
                                          "tasa_de_homicidios", 
                                          "violencia_intrafamiliar")
dim(specification4)
```

## Section 5: Initialize to Git

For the R Markdown version of this document, click on this [LINK](https://github.com/r-rosenberg/GuatemalaProject) to access its repository on GitHub - "GuatemalaProject." 





